---
title: 'Practical 0: Reproducible research in R'
author: "Claire Thiessen"
date: "2025-05-14"
output:
  word_document: 
    fig_height: 6
  pdf_document: default
  html_document:
    fig_height: 7
    highlight: espresso
    theme: readable
---

#loading required packages into R
```{r}
library(tidyverse) 
library(datasauRus)
```

**1. Based on the help file, how many rows and how many columns does the `datasaurus_dozen` file have? What are the variables included in the data frame? (this can be hardcoded)**

run "?datasaurus_dozen" in the Console for info
#answer to 1:
rows: 1846
columns: 3
variables: dataset, x, y

#making a make a *frequency table* of the dataset variable:

```{r}
datasaurus_dozen %>%
  count(dataset)
```
**2. Plot `y` vs. `x` for the `dino` dataset. Then, calculate the correlation coefficient between `x` and `y` for this dataset.**

#filtering to just the dino data
```{r}
dino_data <- datasaurus_dozen %>%
  filter(dataset == "dino")
```

#plotting the dino data
```{r fig.fullwidth=TRUE}
ggplot(data = dino_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

#calculating correlation coefficient for dino data
```{r, eval=FALSE}
dino_data %>%
  summarize(r = cor(x, y))
```
#Dino correlation coefficient
r=-0.064


**3. Plot `y` vs. `x` for the `star` dataset. You can (and should) reuse code we introduced above, just replace the dataset name with the desired dataset. Then, calculate the correlation coefficient between `x` and `y` for this dataset. How does this value compare to the `r` of `dino`?**
```{r}
#creating new data frame with just the star data
star_data <- datasaurus_dozen %>%
  filter(dataset == "star")

#plotting the star data points
ggplot(data = star_data, mapping = aes(x = x, y = y)) +
  geom_point()

#calculating correlation coefficient between x and y of the star dataset
star_data %>%
  summarize(r = cor(x, y))
```
#star correlation coefficient
r=-0.063
This correlation coefficient is nearly identical to that of the dino dataset. 


**4. Plot `y` vs. `x` for the `circle` dataset. You can (and should) reuse code we introduced above, just replace the dataset name with the desired dataset. Then, calculate the correlation coefficient between `x` and `y` for this dataset. How does this value compare to the `r` of `dino`?**
```{r}
#creating new data frame with just the circle data
circle_data <- datasaurus_dozen %>%
  filter(dataset == "circle")

#plotting the circle data points
ggplot(data = circle_data, mapping = aes(x = x, y = y)) +
  geom_point()

#calculating correlation coefficient between x and y of the circle dataset
circle_data %>%
  summarize(r = cor(x, y))
```
#Circle correlation coefficient
r=-0.068
This correlation coefficient is also very similar to the dino and star datasets. 


Facet by the dataset variable, placing the plots in a 3 column grid, and don't add a legend.

**5. Finally, let's plot all datasets at once. In order to do this we will make use of facetting.**

```{r}
ggplot(datasaurus_dozen, aes(x = x, y = y, color = dataset))+
  geom_point()+
  facet_wrap(~ dataset, ncol = 3) +
  theme(legend.position = "none")
```

And we can use the `group_by` function to generate all correlation coefficients.

```{r}
datasaurus_dozen %>%
  group_by(dataset) %>%
  summarize(r = cor(x, y))
```


- **Resize your figures:** 

Click on the gear icon in on top of the R Markdown document, and select "Output Options..." in the dropdown menu. In the pop up dialogue box go to the Figures tab and change the height and width of the figures, and hit OK when done. Then, knit your document and see how you like the new sizes. Change and knit again and again until you're happy with the figure sizes. Note that these values get saved in the YAML.

You can also use different figure sizes for different figures. To do so click on the gear icon within the chunk where you want to make a change. Changing the figure sizes added new options to these chunks: `fig.width` and `fig.height`. You can change them by defining different values directly in your R Markdown document as well.

- **Change the look of your report:** 

Once again click on the gear icon in on top of the R Markdown document, and select "Output Options..." in the dropdown menu. In the General tab of the pop up dialogue box try out different syntax highlighting and theme options. Hit OK and knit your document to see how it looks. Play around with these until you're happy with the look.


---

This set of lab exercises have been adapted from [Mine Çetinkaya-Rundel](https://twitter.com/minebocek)'s class [Introduction to Data Science](https://github.com/ids-s1-19) and [PM566](https://uscbiostats.github.io/PM566/)
